SNP Analysis

### 2018-07-16 SNP_Prediction_Analysis.Rmd
...

source('lib/data_func.R')
source('lib/common_plots.R')

plot.sample<-function(segs, mer)

chr.info = get.chr.lengths(file='hg19_info.txt')

#ptdirs = list.dirs('/Volumes/fh/fast/reid_b/collab/Killcoyne/Data/PerPatient', full.names=T, recursive=F)
#patients = basename(ptdirs)

load('~/Data/Ellie/Analysis/5e6_arms_all_logR/model_data.Rdata', verbose = T)
load('~/Data/Ellie/Analysis/5e6_arms_all_logR/all.pt.alpha.Rdata', verbose = T)
fitV = models$`0.9`
l = performance.at.1se$`0.9`$lambda

@@

ggplot(sm, aes(Pathology, freq, group=Endoscopy, fill=factor(Endoscopy))) + facet_grid(~Status) + geom_col(position = position_dodge()) + ylim(0,1) + title('Pathology frequency per timepoint')
```

@@

```{r, echo=F, eval=T}
qcdata$SampleType = 'BE'
qcdata$SampleType[grep('BLD',rownames(qcdata), ignore.case=T)] = 'Blood Normal'
qcdata$SampleType[grep('gastric',rownames(qcdata), ignore.case=T)] = 'Gastric Normal'
head(qcdata)

@@

@@

#### P: Low SCA, Ploidy<=2
`r round(nrow(lowPloidyCN)/nrow(subset(qcdata, Status == 'P' & SampleType == 'BE')), 3)*100`% of PROGRESSOR samples have <5% SCA and an ASCAT ploidy of 2 or less. Most of these show a >90% purity as well.
`r pander(summary(lowPloidyCN$Purity))`

#### P: No SCA
`r round(nrow(noCN)/nrow(subset(qcdata, Status == 'P' & SampleType == 'BE')), 3)*100`% of PROGRESSOR samples have 0 SCA

@@

@@

totalSamples = table(sub('_.*','', qcdata$Samplename))
lowscaSamples = table(sub('_.*','', lowsca$Samplename))

# For those patients with low SCA samples, a large proportion of them show low SCA in most/all of their samples
#plot(round(lowscaSamples/totalSamples[names(lowscaSamples)], 2))

pt = '403'
pfiles = list.files(paste('~/Data/Reid_SNP/PerPatient', pt, sep='/'), full.names = T)
load(grep('ascat.Rdata', pfiles, value=T), verbose=T)

@@

pt = '512'
pfiles = list.files(paste('~/Data/Reid_SNP/PerPatient', pt, sep='/'), full.names = T)
load(grep('ascat.Rdata', pfiles, value=T), verbose=T)

@@

## Per endoscopy take the average or max bin value as an in silico pooling
sample.info$Level = toupper(sample.info$Level)

range(mergedSegs[1,])
range(x)
plot( x )

x = (mergedSegs[1,]-mean(apply(allDf[,1:ncol(mergedSegs)], 1, mean)))/mean(apply(allDf[,1:ncol(mergedSegs)], 1, sd))
  #sd(apply(allDf[,1:ncol(mergedSegs)], 1, sd))

# names = colnames(mergedSegs)
# mergedSegs = t(apply(mergedSegs, 1, scale, center=T))
# colnames(mergedSegs) = names
# 
# names = colnames(mergedArms)
# mergedArms = t(apply(mergedArms, 1, scale, center=T))
# colnames(mergedArms) = names

samples = unique(subset(sample.info, !Level %in% c('BLD','GASTRIC'), select=c("PatientID", "EndoID")))

meanBinSegs = as.data.frame(matrix(ncol=ncol(mergedSegs), nrow=0, dimnames=list(c(), colnames(mergedSegs))))
meanBinArms = as.data.frame(matrix(ncol=ncol(mergedArms), nrow=0, dimnames=list(c(), colnames(mergedArms))))

mean.qcdata = data.frame(matrix(ncol=8, nrow=0, dimnames=list(c(), c('Samplename','Status','SampleType','Timepoint','mean.sca','mean.ploidy','mean.fit','mean.purity'))))

for (i in 1:nrow(samples)) {
  samples[i,]
  print(paste(samples$PatientID[i], '(',i,')'))

  #rows = subset(sample.info, PatientID == samples$PatientID[i] & EndoID == samples$EndoID[i] )$Samplename
  ptendoID = paste(samples$PatientID[i], samples$EndoID[i], sep='_')

  mean.qcdata[i,] = 
    cbind('Samplename'=ptendoID, subset(qcdata, SampleType == 'BE' & grepl(paste(samples$PatientID[i], samples$EndoID[i], sep='.*'), Samplename)) %>% group_by(Status, SampleType, Timepoint) %>%
    dplyr::summarise(mean.sca=mean(SCA.Ratio), mean.ploidy=mean(Ploidy), mean.fit=mean(Goodnessoffit), mean.purity=mean(Purity)))

  rows = grep(paste(samples$PatientID[i], samples$EndoID[i], sep='.*'), rownames(mergedSegs), value=T)

  sharedR = intersect(rows, rownames(mergedSegs))
  if (length(rows) != length(sharedR)) {
      warning(paste("Missing samples for", endo$PatientID))
      rows = sharedR
  }

  #meanBinSegs[ptendoID,] = apply(mergedSegs[rows,,drop=F], 2, mean)
  meanBinSegs[ptendoID,] = apply(mergedSegs[rows,,drop=F], 2, function(x) x[which.max(abs(x))] )
  #meanBinArms[ptendoID,] = apply(mergedArms[rows,,drop=F], 2, mean)
  meanBinArms[ptendoID,] = apply(mergedArms[rows,,drop=F], 2, function(x) x[which.max(abs(x))] )
}

normalSegs = mergedSegs[grep('BLD|gastric',rownames(mergedSegs),ignore.case=T),]
normalArms = mergedArms[grep('BLD|gastric',rownames(mergedArms),ignore.case=T),]

meanBinSegs = as.matrix(meanBinSegs) 
meanBinArms = as.matrix(meanBinArms)

dim(meanBinSegs)

lowSeg = mergedSegs[grep('^403', rownames(mergedSegs), value=T),,drop=F]
p1 = ggplot(melt(lowSeg), aes(Var2, value)) + facet_grid(~Var1, space='free_x', scales='free') + geom_point() +
  theme(axis.text.x=element_blank() ) + labs(title='Low SCA samples', subtitle='patient 403')
lowSeg = meanBinSegs[grep('^403', rownames(meanBinSegs), value=T),,drop=F]
p2 = ggplot(melt(lowSeg), aes(Var2, value)) + facet_grid(~Var1, space='free_x', scales='free') + geom_point() +
  theme(axis.text.x=element_blank() ) + labs(title='Low SCA samples, mean bins', subtitle='patient 403')

grid.arrange(p1,p2)

highSeg = mergedSegs[grep('^512', rownames(mergedSegs), value=T),,drop=F]
p1 = ggplot(melt(highSeg), aes(Var2, value)) + facet_grid(~Var1, space='free_x', scales='free') + geom_point() +
  theme(axis.text.x=element_blank() ) + labs(title='Mixed SCA samples', subtitle='patient 512')
highSeg = meanBinSegs[grep('^512', rownames(meanBinSegs), value=T),,drop=F]
p2 = ggplot(melt(highSeg), aes(Var2, value)) + facet_grid(~Var1, space='free_x', scales='free') + geom_point() +
  theme(axis.text.x=element_blank() ) + labs(title='Mixed SCA samples, mean bins', subtitle='patient 512')

@@

```{r normals, fig.width=6, fig.height=8, eval=F}
nm = melt(normalSegs)
be = melt(mergedSegs[grep('BLD|gastric', rownames(mergedSegs), ignore.case=T, invert=T),])
beM = melt(meanBinSegs[grep('BLD|gastric', rownames(meanBinSegs), ignore.case=T, invert=T),])

lims = range(nm$value, be$value, beM$value)

grid.arrange(
  ggplot(nm, aes(Var2, value, group=Var2)) + geom_point() + labs(title='Blood/gastric') + theme(axis.text.x = element_blank()) + ylim(lims),
  ggplot(be, aes(Var2, value, group=Var2)) + geom_point() + labs(title='BE samples') + theme(axis.text.x = element_blank()) + ylim(lims),
  ggplot(beM, aes(Var2, value, group=Var2)) + geom_point() + labs(title='BE samples, mean per endo') + theme(axis.text.x = element_blank()) + ylim(lims)
)

@@

Using a random normal sample and a known BE progresser sampe

```{r rnormals, echo=F, fig.width=8, fig.height=8, eval=F}
randNormal = sample(rownames(normalSegs), 1)
pt = unlist(strsplit(randNormal, '_'))[1]
segs = subset(segments.list[[pt]], chr %in% c(1:22) & grepl(randNormal, sample))
segs$chr = factor(segs$chr, levels=c(1:22), ordered = T)

lims = range(segs$medLRR, mergedSegs[randNormal,])
grid.arrange(
  ggplot(segs) + facet_grid(~chr, space='free_x', scales='free') + ylim(lims) +
    geom_segment(aes(x=startpos, xend=endpos, y=medLRR, yend=medLRR), size=3) + 
    theme(axis.text.x=element_blank() ) + labs(title='ASCAT median LRR', subtitle=pt),
  ggplot(melt(as.matrix(mergedSegs[randNormal,])), aes(Var1, value)) + geom_point() + ylim(lims) +
    labs(title='CN Adjusted & winsorized') + theme(axis.text.x=element_blank()),
top='Random BLD or gastric normal', bottom=randNormal)

randBE = '512_18762_133R_27.LogR'
pt = unlist(strsplit(randBE, '_'))[1]
segs = subset(segments.list[[pt]], chr %in% c(1:22) & grepl(randBE, sample))
segs$chr = factor(segs$chr, levels=c(1:22), ordered = T)

lims = range(segs$medLRR, mergedSegs[sub('\\.LogR','',randBE),])
grid.arrange(
  ggplot(segs) + facet_grid(~chr, space='free_x', scales='free') + ylim(lims) +
    geom_segment(aes(x=startpos, xend=endpos, y=medLRR, yend=medLRR), size=3) + 
    theme(axis.text.x=element_blank() ) + labs(title='Median LRR, ASCAT segments', subtitle=pt), 
  ggplot(melt(as.matrix(mergedSegs[sub('\\.LogR','',randBE),])), aes(Var1, value)) + geom_point() + ylim(lims) + 
    theme(axis.text.x=element_blank()) + labs(title='Winsorized LRR, binned'),  
top='BE', bottom=randBE)

@@

Using the mean bin values.
```{r, warning=F}
adjust.cols<-function(mtx, means=NULL, sds=NULL) {
  if (!is.null(means)) {
    if (ncol(mtx) != length(means))
      stop("Vector of means needs to match the columns in the matrix")
    for (i in 1:ncol(mtx)) 
      mtx[,i] = unit.var(mtx[,i], means[i], sds[i])
  } else {
    for (i in 1:ncol(mtx)) 
      mtx[,i] = unit.var(mtx[,i])
  }
  return(mtx)
}

ns = adjust.cols(normalSegs, z.mean, z.sd)
na = adjust.cols(normalArms, z.arms.mean, z.arms.sd)

ms = adjust.cols(meanBinSegs, z.mean, z.sd)
ms = adjust.cols(mergedSegs, z.mean, z.sd)
range(ms)
dim(ms)

ma = adjust.cols(meanBinArms, z.arms.mean, z.arms.sd)
ma = adjust.cols(mergedArms, z.arms.mean, z.arms.sd)
range(ma)
dim(ma)

nm = melt(ns)
be = melt(ms[grep('BLD|gastric', rownames(ms), ignore.case=T, invert=T),])

@@

@@

```{r echo=F}
cx = score.cx(ms, 1)
arrayDf = subtract.arms(ms, ma)
arrayDf = cbind(arrayDf, 'cx'=unit.var(cx, mn.cx, sd.cx))

cxN = score.cx(ns, 1)
arrayDfNorm = subtract.arms(ns, na)
arrayDfNorm = cbind(arrayDfNorm, 'cx'=unit.var(cxN, mn.cx, sd.cx))

@@

@@

@@

```{r, warning=F, fig.height=8}
prob = predict(fitV, newx=arrayDf, s=l, type='response')
rr = predict(fitV, newx=arrayDf, s=l, type='link')

preds = cbind.data.frame(rownames(prob),prob, rr)
colnames(preds) = c('Samplename','Prob', 'RR')

#offset = log(0.0225)
# preds = preds %>% dplyr::mutate( 
#   'Adj. RR'=RR+offset,  
#   'Adj. Prob'=1/(1+exp(-RR+abs(offset)) )
# )

@@

A quick look at the two random samples from earlier.  
`r pander(subset(preds, Samplename == randNormal), caption='Random normal')`

The random BE shows a very high probability:
`r pander(subset(preds, Samplename == randBE), caption='Random BE')`

@@

```{r, warning=F, fig.height=8}
# normals = subset(preds, grepl('BLD|gastric', Samplename, ignore.case=T))
# grid.arrange(
#   ggplot(normals, aes(Prob)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
#     scale_fill_gradientn(colors = myPal,  name='') + 
#     labs(title='Predictions on BLD/Gastric', y='n Samples', x='Unadjusted Probability'),
# top='BLD/gastric normals' )
# 
# normals[which(normals$Prob > 0.5),]

pred.qc = base::merge(preds, mean.qcdata, by='Samplename')
#pred.qc = base::merge(preds, qcdata, by='Samplename')

prog = subset(pred.qc, !grepl('BLD|gastric', Samplename, ignore.case=T) & Status == 'P')
ggplot(prog, aes(Prob)) + facet_grid(~Timepoint) + 
  geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
  scale_fill_gradientn(colors = myPal,  name='') + 
  labs(title='Predictions on progressors', y='n Samples', x='Unadjusted Probability')

nonprog = subset(pred.qc, !grepl('BLD|gastric', Samplename, ignore.case=T) & Status == 'NP')
ggplot(nonprog, aes(Prob)) + facet_grid(~Timepoint) + 
  geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
  scale_fill_gradientn(colors = myPal,  name='') + 
  labs(title='Predictions on non-progressors', y='n Samples', x='Unadjusted Probability')

@@

```{r roc, echo=T}
si = do.call(rbind.data.frame,c(strsplit(as.character(pred.qc$Samplename), '_'), stringsAsFactors=F) )
if (ncol(si) == 4) {
  colnames(si) = colnames(sample.info)[1:4]
} else {
  colnames(si) = colnames(sample.info)[1:2]
}
pred.qc = cbind(si, pred.qc)
pred.qc = base::merge( pred.qc, patient.info[,c('PatientID','Pathology')], by='PatientID')

rows = grep('BLD|gastric', pred.qc$Samplename, ignore.case=T)
if (length(rows) > 0) {
  normals = pred.qc[rows,]
  pred.qc = pred.qc[-rows,]
}
nrow(pred.qc)

@@

# Exclude those odd tetraploidy NPs
pred.qc[which(with(pred.qc, (Samplename %in% highPloidyCN$Samplename))), 'exclude'] = T
pred.qc[which(pred.qc$PatientID %in% subset(pred.qc, Samplename %in% highPloidyCN$Samplename)$PatientID), 'exclude'] = T

# Exclude HGD NP patients
pred = subset(pred.qc, !exclude)

roc = pROC::roc(Status ~ Prob, data=pred, auc=T, ci=T, of='thresholds')
roc.plot(roc)

#rocNoLow = pROC::roc(Status ~ Prob, data=subset(pred, !Samplename %in% c(lowsca$Samplename)), auc=T, ci=T, of='thresholds')
#roc.plot(rocNoLow)

per_pt = pred %>% group_by(PatientID, Timepoint, Status) %>% dplyr::summarise( max.Prob = max(Prob), max.RR = max(RR) )
rocpp = pROC::roc(Status ~ max.Prob, data=per_pt, auc=T, ci=T, of='thresholds')
roc.plot(rocpp)

per_pt_ex = subset(pred, !Samplename %in% c(lowsca$Samplename,highPloidyCN$Samplename)) %>% group_by(PatientID, Timepoint, Status) %>% dplyr::summarise(
  max.Prob = max(Prob), max.RR = max(RR)
)
rocppex = pROC::roc(Status ~ max.Prob, data=per_pt_ex, auc=T, ci=T, of='thresholds')
roc.plot(rocppex)

roct2 = pROC::roc(Status ~ Prob, data=subset(pred, Timepoint != 'T1' & !Samplename %in% c(lowsca$Samplename)), auc=T, ci=T, of='thresholds')
roc.plot(roct2)

>>>>

suppressPackageStartupMessages(source('lib/data_func.R'))
suppressPackageStartupMessages(source('lib/common_plots.R'))

chr.info = get.chr.lengths(file='~/tmp/hg19_info.txt')

adjust.cols<-function(mtx, means=NULL, sds=NULL) {
  if (!is.null(means)) {
    if (ncol(mtx) != length(means))
      stop("Vector of means needs to match the columns in the matrix")
    for (i in 1:ncol(mtx)) 
      mtx[,i] = unit.var(mtx[,i], means[i], sds[i])
  } else {
    for (i in 1:ncol(mtx)) 
      mtx[,i] = unit.var(mtx[,i])
  }
  return(mtx)
}

load('~/Data/Ellie/Analysis/5e6_arms_all_logR/model_data.Rdata', verbose = T)
load('~/Data/Ellie/Analysis/5e6_arms_all_logR/all.pt.alpha.Rdata', verbose = T)
#alpha = '0'
alpha = '0.9'
fitV = models[[alpha]]
l = performance.at.1se[[alpha]]$lambda

@@

ggplot(sm, aes(Pathology, freq, group=Endoscopy, fill=factor(Endoscopy))) + facet_grid(~Status) + 
  geom_col(position = position_dodge()) + ylim(0,1) + labs(title='Pathology frequency per timepoint')
```

@@

```{r, echo=F, warning=F}
qcdata$SampleType = 'BE'
qcdata$SampleType[grep('BLD',rownames(qcdata), ignore.case=T)] = 'Blood Normal'
qcdata$SampleType[grep('gastric',rownames(qcdata), ignore.case=T)] = 'Gastric Normal'
head(qcdata)

@@

## PER ENDOSCOPY
qcdata.samples = qcdata %>% group_by(Samplename) %>% dplyr::mutate(
  PatientID = unlist(strsplit(Samplename, '_'))[1],
  EndoID = unlist(strsplit(Samplename, '_'))[3],
  Level = unlist(strsplit(Samplename, '_'))[4] 
)

qcdata = qcdata.samples %>% group_by(PatientID, EndoID, Status, Timepoint, SampleType) %>% dplyr::summarise(
  Ploidy = mean(Ploidy), Purity = mean(Purity), Goodnessoffit = mean(Goodnessoffit), SCA.Ratio = mean(SCA.Ratio)
)

@@

#### P: Low SCA, Ploidy<=2
`r round(nrow(lowPloidyCN)/nrow(subset(qcdata, Status == 'P' & SampleType == 'BE')), 3)*100`% of PROGRESSOR endoscopies have <5% SCA and an ASCAT ploidy of 2 or less. Most of these show a >90% purity as well.
`r pander(summary(lowPloidyCN$Purity))`

#### P: No SCA
`r round(nrow(noCN)/nrow(subset(qcdata, Status == 'P' & SampleType == 'BE')), 3)*100`% of PROGRESSOR endoscopies have 0 SCA

@@

```{r, echo=F, warning=F}

@@

pt = '403'
pfiles = list.files(paste('~/Data/Reid_SNP/PerPatient', pt, sep='/'), full.names = T)
load(grep('ascat.Rdata', pfiles, value=T), verbose=F)

@@

pt = '512'
pfiles = list.files(paste('~/Data/Reid_SNP/PerPatient', pt, sep='/'), full.names = T)
load(grep('ascat.Rdata', pfiles, value=T), verbose=F)

@@

dim(mergedSegs)
dim(mergedArms)

nm.rows = grep('BLD|GASTRIC', rownames(mergedSegs))
normalSegs = mergedSegs[nm.rows,]
normalArms = mergedArms[nm.rows,]

nrow(normalSegs) == nrow(normalArms)

mergedSegs = mergedSegs[-nm.rows,]
mergedArms = mergedArms[-nm.rows,]

nrow(mergedSegs) == nrow(mergedArms)

lowSeg = mergedSegs[grep('^403', rownames(mergedSegs), value=T),,drop=F]
p1 = ggplot(melt(lowSeg), aes(Var2, value)) + facet_grid(~Var1, space='free_x', scales='free') + geom_point() +
  theme(axis.text.x=element_blank() ) + labs(title='Low SCA samples', subtitle='patient 403')

highSeg = mergedSegs[grep('^512', rownames(mergedSegs), value=T),,drop=F]
p2 = ggplot(melt(highSeg), aes(Var2, value)) + facet_grid(~Var1, space='free_x', scales='free') + geom_point() +
  theme(axis.text.x=element_blank() ) + labs(title='Mixed SCA samples', subtitle='patient 512')

@@

```{r normals, fig.width=6, fig.height=8, eval=F}
nm = melt(normalSegs)
be = melt(mergedSegs[-nm.rows,])

lims = range(nm$value, be$value)

grid.arrange(
  ggplot(nm, aes(Var2, value, group=Var2)) + geom_point() + labs(title='Blood/gastric') + theme(axis.text.x = element_blank()) + ylim(lims),
  ggplot(be, aes(Var2, value, group=Var2)) + geom_point() + labs(title='BE samples') + theme(axis.text.x = element_blank()) + ylim(lims)
)

@@

Using a random normal sample and a known BE progresser sample

```{r rnormals, echo=F, fig.width=8, fig.height=8, eval=T}
randNormal = sample(rownames(normalSegs), 1)
pt = unlist(strsplit(randNormal, '_'))[1]
segs = subset(segments.list[[pt]], chr %in% c(1:22) & grepl(sub('_', '.*', randNormal),sample))
segs$chr = factor(segs$chr, levels=c(1:22), ordered = T)

lims = range(segs$medLRR, normalSegs[randNormal,])
grid.arrange(
  ggplot(segs) + facet_grid(~chr, space='free_x', scales='free') + ylim(lims) +
    geom_segment(aes(x=startpos, xend=endpos, y=medLRR, yend=medLRR), size=3) + 
    theme(axis.text.x=element_blank() ) + labs(title='ASCAT median LRR', subtitle=pt),
  ggplot(melt(as.matrix(normalSegs[randNormal,])), aes(Var1, value)) + geom_point() + ylim(lims) +
    labs(title='CN Adjusted & winsorized') + theme(axis.text.x=element_blank()),
top='Random BLD or gastric normal', bottom=randNormal)

randBE = '512_133R'
randBE = '1006_190V'
pt = unlist(strsplit(randBE, '_'))[1]
segs = subset(segments.list[[pt]], chr %in% c(1:22) & grepl(sub('_', '.*',randBE), sample))
segs$chr = factor(segs$chr, levels=c(1:22), ordered = T)

lims = range(segs$medLRR, mergedSegs[sub('\\.LogR','',randBE),])
grid.arrange(
  ggplot(segs) + facet_grid(~chr, space='free_x', scales='free')  +
    geom_segment(aes(x=startpos, xend=endpos, y=medLRR, yend=medLRR), size=3) + 
    theme(axis.text.x=element_blank() ) + labs(title='Median LRR, ASCAT segments', subtitle=pt), 
  ggplot(melt(as.matrix(mergedSegs[sub('\\.LogR','',randBE),])), aes(Var1, value)) + geom_point() +
    theme(axis.text.x=element_blank()) + labs(title='Winsorized LRR, binned'),  
  ggplot(melt(adjust.cols(mergedSegs[sub('\\.LogR','',randBE),,drop=F], z.mean, z.sd)), aes(Var2, value)) + geom_point() +
    theme(axis.text.x=element_blank()) + labs(title='Col adjusted LRR, binned'),  
top='BE', bottom=randBE)

sa = subtract.arms(adjust.cols(mergedSegs[sub('\\.LogR','',randBE),,drop=F], z.mean, z.sd),
              adjust.cols(mergedArms[sub('\\.LogR','',randBE),,drop=F], z.arms.mean, z.arms.sd))[,1:589]
ggplot(melt(as.matrix(sa)), aes(Var1, value)) + geom_point()

@@

Using the mean bin values (merged during tiling).
```{r, warning=F}
# z.mean[is.na(z.mean)] = median(z.mean,na.rm=T)
# z.arms.mean[is.na(z.arms.mean)] = median(z.arms.mean,na.rm=T)
# 
# z.sd[is.na(z.sd)] = median(z.sd,na.rm=T)
# z.arms.sd[is.na(z.arms.sd)] = median(z.arms.sd,na.rm=T)

ns = adjust.cols(normalSegs, z.mean, z.sd)
#ns = apply(normalSegs, 2, unit.var)
range(ns)

na = adjust.cols(normalArms, z.arms.mean, z.arms.sd)
#na = apply(normalArms, 2, unit.var)
range(na)

ms = adjust.cols(mergedSegs, z.mean, z.sd)
#ms = apply(mergedSegs, 2, unit.var)
range(ms)
dim(ms)

ma = adjust.cols(mergedArms, z.arms.mean, z.arms.sd)
#ma = apply(mergedArms, 2, unit.var)
range(ma)
dim(ma)

nm = melt(ns)
be = melt(ms)

@@

grid.arrange(
  ggplot(melt(ms[sub('\\.LogR','',randBE),,drop=F]), aes(Var2, value)) + geom_point() +
    theme(axis.text.x=element_blank()) + labs(title='Col adjusted LRR, binned'),  
  ggplot(melt(subtract.arms(ms[sub('\\.LogR','',randBE),,drop=F],ma[sub('\\.LogR','',randBE),,drop=F])), aes(Var2, value)) +
    geom_point() + theme(axis.text.x=element_blank()) + labs(title='Arms subtracted col adjusted LRR, binned'),  
  top=randBE)
# So what if I don't subtract the arms?

@@

```{r echo=F}
cxN = score.cx(ns, 1)
arrayDfNorm = subtract.arms(ns, na)
#arrayDfNorm = cbind(ns, na)
arrayDfNorm = cbind(arrayDfNorm, 'cx'=unit.var(cxN, mn.cx, sd.cx))

@@

preds.normal = cbind.data.frame(rownames(probN),probN, predict(fitV, newx=arrayDfNorm, s=l, type='link'))
colnames(preds.normal) = c('Samplename','Prob', 'RR')

cx = score.cx(ms, 1)
#arrayDf = cbind(ms,ma)
arrayDf = subtract.arms(ms, ma)

arrayDf[,589:ncol(arrayDf)] = arrayDf[,589:ncol(arrayDf)]+abs(mean(arrayDf[,589:ncol(arrayDf)]))

arrayDf = cbind(arrayDf, 'cx'=unit.var(cx, mn.cx, sd.cx))

grid.arrange(
  ggplot(melt(arrayDf), aes(Var2, value)) + geom_point() + 
    geom_point(data=melt(arrayDf[,589:634]), col='red') + labs(title='SNP'),
  ggplot(melt(allDf), aes(Var2, value)) + geom_point() +
      geom_point(data=melt(allDf[,589:634]), col='red') + labs(title='sWGS'),
nrow=1)

# grid.arrange(
#     ggplot(melt(arrayDf), aes(Var2, value)) + geom_point() + geom_point(data=melt(arrayDf[,589:634]+abs(mean(arrayDf[,589:634])*2)), col='red') + labs(title='SNP'),
#     ggplot(melt(allDf), aes(Var2, value)) + geom_point() +
#       geom_point(data=melt(allDf[,589:634]), col='red') + labs(title='sWGS'),
#   nrow=1)

prob = predict(fitV, newx=arrayDf, s=l, type='response')
rr = predict(fitV, newx=arrayDf, s=l, type='link')

preds = cbind.data.frame(rownames(prob),prob, rr, stringsAsFactors=F)
colnames(preds) = c('Samplename','Prob', 'RR')

## TEMP: I accidentally included normals by failing to account for difference in name
#preds = preds[-grep('_\\d+$', rownames(preds)),]

@@

```{r, eval=F}
x = apply(mergedSegs, 1, function(x) {
  ds = c()
  for (i in 1:(length(x)-1)) {
    ds = c(ds, abs(x[[i]]) - abs(x[[i+1]]))
  }
  abs(diff(range(ds)))
})
plot(x, col=pred.qc$Status)

x1 = apply(ms, 1, function(x) {
  ds = c()
  for (i in 1:(length(x)-1)) {
    ds = c(ds, abs(x[[i]]) - abs(x[[i+1]]))
  }
  abs(diff(range(ds)))
})
plot(x1, col=pred.qc$Status)

y = apply(normalSegs, 1, function(x) {
  ds = c()
  for (i in 1:(length(x)-1)) {
    ds = c(ds, abs(x[[i]]) - abs(x[[i+1]]))
  }
  abs(diff(range(ds)))
})
plot(y)

range(y)
range(x)

nprgs = subset(pred.qc, Status == 'NP')
prgs = subset(pred.qc, Status == 'P')

t.test(x[prgs$Samplename],x[nprgs$Samplename]) 
t.test(x[prgs$Samplename],y)
t.test(x[nprgs$Samplename],y)
```

@@

```{r, warning=F, fig.height=8}
#offset = log(0.0225)
# preds = preds %>% dplyr::mutate( 
#   'Adj. RR'=RR+offset,  
#   'Adj. Prob'=1/(1+exp(-RR+abs(offset)) )
# )

@@

A quick look at the two random samples from earlier.  
`r pander(subset(preds.normal, Samplename == randNormal), caption='Random normal')`

The random BE should show a very high probability, but for some reason is missed entirely:
`r pander(subset(preds, Samplename == randBE), caption='Random BE')`

@@

```{r, warning=F, fig.height=8}
ggplot(preds.normal, aes(Prob)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
  scale_fill_gradientn(colors = myPal,  name='') +
  labs(title='Predictions on BLD/Gastric', y='n Samples', x='Unadjusted Probability')

qcdata.be = subset(qcdata, SampleType == 'BE')

get.qcinfo<-function(vec) {
  endo = vec[2]
  pt = vec[1]
  y = subset(qcdata.be, PatientID == pt & EndoID == endo)
  y = y %>% group_by(Status, Timepoint) %>%  dplyr::summarise( mean(Ploidy), mean(Purity), mean(Goodnessoffit), mean(SCA.Ratio) )
  cbind.data.frame('PatientID'=pt, 'EndoID'=endo, y, stringsAsFactors = F)
}

qc = do.call(rbind.data.frame, c( lapply(preds$Samplename, function(x) get.qcinfo( unlist(strsplit(x, '_'))) ), stringsAsFactors=F))

pred.qc = cbind.data.frame(preds, qc)
dim(pred.qc)
table(pred.qc$Status)

ggplot(pred.qc, aes(`mean(SCA.Ratio)`, Prob, color=Status)) + geom_point()
ggplot(pred.qc, aes(`mean(Ploidy)`, Prob, color=Status)) + geom_point()

pander(cor.test(pred.qc$`mean(SCA.Ratio)`, pred.qc$Prob))

prog = subset(pred.qc, Status == 'P')
ggplot(prog, aes(Prob)) + facet_grid(~Timepoint) + 
  geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
  scale_fill_gradientn(colors = myPal,  name='') + 
  labs(title='Predictions on progressors', y='n Samples', x='Unadjusted Probability')

nonprog = subset(pred.qc, Status == 'NP')
ggplot(nonprog, aes(Prob)) + facet_grid(~Timepoint) + 
  geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
  scale_fill_gradientn(colors = myPal,  name='') + 
  labs(title='Predictions on non-progressors', y='n Samples', x='Unadjusted Probability')

@@

Using my thresholds

```{r roc, echo=T}
pred.qc = base::merge( pred.qc, patient.info[,c('PatientID', 'Timepoint Code','Pathology')], by.x=c('PatientID', 'EndoID'), by.y=c('PatientID','Timepoint Code') )

@@

# pred.qc.ex = subset(pred.qc, !exclude)
# cuts = seq(0,1,0.1)
# 
# pred.qc.ex$quants = with(pred.qc.ex, cut(Prob, breaks=cuts))
# qt = as.data.frame.matrix(table(pred.qc.ex$quants, pred.qc.ex$Status))
# qt$quant = rownames(qt)

get.risk<-function(prob) {
  if (prob < 0.3) return('Low')
  if (prob >= 0.3 & prob < 0.7) return('Moderate')
  if (prob >= 0.7) return('High')
}

# All data
pred.qc = pred.qc %>% dplyr::rowwise() %>% dplyr::mutate(
  Risk = get.risk(Prob)
)
pred.qc$Risk = factor(pred.qc$Risk, levels=c('Low','Moderate','High'), ordered = T)

roc = pROC::roc(Status ~ Prob, data=pred.qc, auc=T, ci=T, of='thresholds')
roc.plot(roc) + labs(title='All, Prob')
pander(pROC::coords(roc, 'best'), caption='All samples, ROC best')

roc = pROC::roc(Status ~ Risk, data=pred.qc, auc=T, ci=T, of='thresholds')
roc.plot(roc) + labs(title='All, Risk')

# Exclude those odd tetraploidy NPs
#pred.qc[which(with(pred.qc, (Samplename %in% highPloidyCN$Samplename))), 'exclude'] = T
pred.qc[which(pred.qc$PatientID %in% highPloidyCN$PatientID & pred.qc$EndoID %in% highPloidyCN$EndoID), 'exclude'] = T

# Exclude HGD NP patients
roc = pROC::roc(Status ~ Prob, data=subset(pred.qc, !exclude), auc=T, ci=T, of='thresholds')
roc.plot(roc) + labs(title='Excluded low sca/NP tetra), Prob')

roc = pROC::roc(Status ~ Risk, data=subset(pred.qc, !exclude), auc=T, ci=T, of='thresholds')
roc.plot(roc) + labs(title='Excluded low sca/NP tetra), Risk')

per_pt = 
  subset(pred.qc, !exclude) %>% group_by(PatientID, Status) %>% dplyr::summarise( max.Prob = max(Prob), max.RR = max(RR), max.Risk = max(Risk) )
rocpp = pROC::roc(Status ~ max.Prob, data=per_pt, auc=T, ci=T, of='thresholds')
roc.plot(rocpp) + labs(title='Max per patient (excluded low sca/NP tetra), Prob')

rocpp = pROC::roc(Status ~ max.Risk, data=per_pt, auc=T, ci=T, of='thresholds')
roc.plot(rocpp) + labs(title='Max per patient (excluded low sca/NP tetra), Risk')

# Exclude the low SCA patients
pred.qc[which(pred.qc$PatientID %in% lowsca$PatientID & pred.qc$EndoID %in% lowsca$EndoID), 'exclude'] = T

per_pt_ex = subset(pred.qc, !exclude) %>% group_by(PatientID, Timepoint, Status) %>% dplyr::summarise(
  max.Prob = max(Prob), max.RR = max(RR), max.Risk = max(Risk)
)
rocppex = pROC::roc(Status ~ max.Prob, data=per_pt_ex, auc=T, ci=T, of='thresholds')
roc.plot(rocppex) + labs(title='Max per patient, Prob')

rocppex = pROC::roc(Status ~ max.Risk, data=per_pt_ex, auc=T, ci=T, of='thresholds')
roc.plot(rocppex) + labs(title='Max per patient, Risk')

roct2 = pROC::roc(Status ~ Prob, data=subset(pred.qc, Timepoint != 'T1' & !exclude), auc=T, ci=T, of='thresholds')
roc.plot(roct2) + labs(title='T2, Prob')

roct2 = pROC::roc(Status ~ Risk, data=subset(pred.qc, Timepoint != 'T1' & !exclude), auc=T, ci=T, of='thresholds')
roc.plot(roct2) + labs(title='T2, Risk')

ggplot(subset(pred.qc, !exclude), aes(`mean(SCA.Ratio)`,Prob, color=Risk, shape=Status)) + 
  facet_grid(~Status) + geom_point()
